---
title: "Overview of basic features of ggplot"
author: "Caleb Trujillo"
date: "1/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require("tidyverse")
require("nycflights13")
require("gapminder")
require("Lahman")
```

##  Scatter Plot

Data can be visualized from a data table to show two continuous numerical variables as a scatter plot such in the graph below. The x-axis shows `displ`, the size of a car’s engine in liters and the y-axis shows `hwy` in miles per gallon (mpg) to indicate each car’s fuel efficiency on the highway. The graph also shows the classification of the vehicle by `class`, indicated by color and shape.

```{r scatter plot, echo = FALSE}

ggplot(data = mpg, mapping = aes()) +
  geom_point(mapping = aes())+
  geom_smooth()

```

The graph displays a trend of smaller classes of cars tend to have better gas millage. 

## Bar Plot

In addition to viewing data as a scatter plot, you can also visualize categorical data by counts or relative percentages.  

```{r bar plot, echo = FALSE}

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")

```

## Box plot

You can also embed plots, for example:

```{r box plot,  echo = FALSE}

```


